Determining sequence-dependent DNA oligonucleotide hybridization and dehybridization mechanisms using coarse-grained molecular simulation, Markov state models, and infrared spectroscopy

### 2021-07-31 manuscript_marked_up.tex
Andy edits

Over the last couple of decades, DNA has proven to be much more than a vessel for genetic information. From sensing to computing to directed self-assembly, the programmable and predictable nature of DNA has unlocked numerous unforeseen nanotechnology applications.\citep{Seeman2017DNANanotechnology, Adleman1994MolecularProblems, Rothemund2006FoldingPatterns, Gu2010ALine} Recently, single molecule localization techniques have exploited the rapid and transient binding of short DNA oligomers in order to achieve super-resolution microscopy and optical multiplexing. \cite{Schnitzbauer2017Super-resolutionDNA-PAINT, Shah2019, Strauss2020UpDNA-PAINT} Predictive understanding of the sequence-dependent thermodynamics of DNA hybridization/dehybridization -- the assembly/disassembly of a DNA duplex from two single strands -- has underpinned the rational design of DNA oligomer sequences for nanotechnology applications, where sequence-dependent nearest-neighbor models can accurately account for mismatched pairs, dangling ends, and other non-native bonding effects.\citep{SantaLucia1998AThermodynamics, Santalucia2004TheMotifs} Secondary DNA structures such as hairpins and G-quadruplexes have also been studied in depth and leveraged for nanotechnology applications.\citep{Tsukanov2013DetailedOrigami, Mosayebi2014TheFormation, Mergny2019DNANanotechnology} Predictive models of the dynamical, as opposed to purely thermodynamical, behaviors of DNA have become increasingly important in developing technologies such as DNA-PAINT (DNA Points Accumulation For Imaging In Nanoscale Topography), but these technologies have outpaced our fundamental understanding of the dynamics themselves. \citep{Deluca2020DynamicDevices, Cordes2010SensingTransfer, Naimark2020DNADevelopment} Many experimental and computational studies have investigated DNA dynamical phenomena over picosecond to millisecond time scales.\citep{Yin2011KineticsHybridization, Xiao2019, Hinckley2014Coarse-grainedEffects, Sanstead2016, Porschke1971CooperativeTransition} Kinetic models have been developed for particular DNA processes such as toehold exchanges and optical barcoding \citep{Zhang2009ControlExchange, Shah2019ProgrammingFingerprinting} and supervised machine learning techniques have been combined with experimental measurements to predict the on/off rates as a function of sequence. \citep{Schickinger2018TetheredHelices, Zhang2018PredictingSequence, Shah2019}. A comprehensive understanding of the full dynamical landscape of hybridization/dehybridization accounting for the sequence-dependent metastable states and association/dissociation pathways remains lacking and fundamental questions remain unresolved. For example, it remains unclear the extent to which hybridization of short DNA oligomers largely proceeds in a conventionally assumed ``all-or-nothing'' fashion or if long-lived metastable states facilitate the transition.\citep{Araque2016LatticeCooperativity, Sikora2013ModelingIntermediates, Wyer2014KineticsAT-tracts, Sanstead2016, Sanstead2018DirectDehybridization, Xiao2019} Out-of-register ``shifted'' base paired structures \citep{Hinckley2014Coarse-grainedEffects, Maciejczyk2014DNAModel, Araque2016LatticeCooperativity, Xiao2019} and frayed structures \citep{Zgarbova2014BaseRNA, Nonin1995TerminalFraying, Nikolova2012ProbingSimulations, Andreatta2006UltrafastHelix} stand as candidates for metastable states with the potential to mediate substantial deviations from all-or-nothing behavior, but the degree to which these states are kinetically relevant is difficult to determine experimentally and is likely to be highly sequence-dependent. The development of predictive models and design rules with which to engineer DNA strands with tailored hybridization/dehybridization kinetics and pathways is vital to advancing rational design of DNA strands for nanotechnology applications and is also of importance in understanding fundamental biological processes such as transcription and gene regulation.

Our understanding of hybridization dynamics is built upon decades of experiments -- such as temperature-jump, salt-jump, pH-jump, and other perturbative methods -- that rapidly stimulate DNA and monitor relaxation to a new equilibrium.\citep{Morrison1993SensitiveSolution, Wetmur1968KineticsDNA, Craig1971RelaxationOligonucleotides, Porschke1973ThermodynamicsPairs, Williams1989LaserDGCATGC, Narayanan2012ExploringMixing, Chen2007InfluenceHybridization, Sanstead2018DirectDehybridization} More recently, single molecule diffusion and tethered multifluorophore assays have facilitated analyses under equilibrium conditions, but these results can be hampered by slow data collection rates and fluorescent tags effects on strand dynamics, particularly for shorter oligomers.\citep{Liu20173DSolution,  Schickinger2018TetheredHelices, Chen2008Base-by-baseSpectroscopy, Dupuis2013Single-moleculeHelices} A number of computational modeling approaches have also been employed to provide molecular-level resolution of hybridization. Simplified lattice models can recapitulate the essential aspects of the hybridization pathways but lack the realism of continuous space representations \citep{Araque2016LatticeCooperativity, Phys2019}. The long time scales associated with hybridization/dehybridization events place them outside the reach of unbiased all-atom molecular dynamics simulations,\citep{Hinckley2014Coarse-grainedEffects} but they can be observed by employing enhanced sampling techniques\citep{Piana2007AtomisticTransition, Zerze2021ThermodynamicsSimulations, Hinckley2013AnHybridization, Schmitt2013ExploringSurface, Sambriski2009, Sambriski2009SequencePathways, Hoefert2011MolecularOligonucleotides,Romano2013DNADependence, Hinckley2014Coarse-grainedEffects} or by using elevated temperature or denaturing solvent concentrations to induce one-way dissociation events.\citep{Wong2008TheSimulations, Perez2010Real-timeUnfolding} The effect of the applied bias upon the thermodynamics can be rigorously corrected for using standard reweighting techniques \citep{Gallicchio2005TemperaturePaths, Souaille2001ExtensionCalculations, Shirts2008StatisticallyStates, Kumar1992THEMethod}. \rood{Rigorous elimination of the bias in the kinetics is critical for the construction of robust and reliable Markov state models that reflect the true system dynamics and are uncontaminated by any residual effects of the biasing potentials used to induce good sampling. A number of approaches to correct the kinetics are also available, including Girsanov reweighting, transition-based reweighting analysis (TRAM), dynamic histogram analysis method (DHAM), and their derivatives~\citep{Prinz2011OptimalDynamics, Chodera2011DynamicalTemperatures, Stelzl2017DynamicSimulations, Donati2017GirsanovModels, Donati2018GirsanovSimulations, Quer2018ANCOORDINATES} \citep{Mey2014XTRAM:States}. The application of these methods under the conditions of high bias necessary for good sampling can, however, present challenges for numerical convergence.} A number of coarse-grained DNA force fields have been developed that enable direct observation of these events over microsecond time scales via unbiased coarse-grained molecular dynamics simulations, \citep{Romano2013DNADependence, Hinckley2013AnHybridization, Maciejczyk2014DNAModel, Markegard2015, Dans2016MultiscaleDNA} which, up to an acceleration factor associated with the smoothing of the underlying free energy landscape inherent to the coarse-graining procedure, can preserve a faithful model of the unbiased dynamics and associated pathways. These models have previously been used to study biological phenomena such as nucleosome dynamics\citep{Lequieu2016Tension-dependentUnwrapping, Lequieu2017InSliding} and transcription factor binding \cite{Terakawa2015P53Simulations, Tan2018DynamicData} as well as nanotechnology applications such as strand displacement \citep{Srinivas2013OnDisplacement, Haley2020DesignDisplacement} and DNA origami \citep{Snodin2019Coarse-grainedOrigami, Doye2020TheOrigami}. \rood{In this work, we choose to employ a coarse-grained model for DNA \citep{Hinckley2013AnHybridization} that is sufficiently inexpensive to enable the collection of sufficient volumes of unbiased simulation trajectories and adequately sample configurational space that we do not need to appeal to biasing strategies to enhance convergence nor apply any \textit{post hoc} corrections to the thermodynamics or kinetics.}

@@

We first sought to demonstrate that the coarse-grained 3SPN.2 model accurately recapitulates the experimentally observed kinetics of DNA oligomer hybridization/dissociation by validating our computational predictions against temperature-jump infrared (T-jump IR) experiments. We conducted T-jump IR experiments for each DNA sequence as a function of temperature and extracted \rood{the ``slow'' $k_d^\mathrm{slow}$  and ``fast'' $k_d^\mathrm{fast}$ rates, corresponding to processes on the 10-30 $\mu$s and 70-100 ns timescales, respectively}. The slow response has previously been attributed to duplex dissociation on microsecond time scales induced by the rapid heating of the initially hybridized duplex.~\citep{Sanstead2016, Sanstead2018DirectDehybridization} The fast response has been assigned to terminal base pair fraying,~\citep{Sanstead2016, Sanstead2018DirectDehybridization} with this process corresponding to a relatively complex dynamical process that can span time scales from picoseconds to microseconds.~\cite{Nonin1995TerminalFraying, Nikolova2012ProbingSimulations, Andreatta2006UltrafastHelix, Galindo-Murillo2015ConvergenceDGCACGAACGAACGAACGC} All-atom simulations suggest that frayed ends can assume misaligned WC bonds, base-sugar hydrogen bonds, and terminal stacked conformations.~\citep{Pinamonti2019TheModels, Zgarbova2014BaseRNA}

To computationally mimic the T-jump process in our 3SPN.2 simulations, we conducted 1 $\mu$s simulations of an initially hybridized DNA duplex over a range of temperatures and monitored its structural relaxation. We performed 120 independent simulations for each DNA sequence at each temperature, and from these extracted computational estimates of $k_d^\mathrm{slow}$ and $k_d^\mathrm{fast}$ (\blauw{Fig.~S4}). First, we tracked the slow response corresponding to duplex dissociation in our simulations by compiling the distribution of times at which both of the central base pairs first separate to a distance of 1.3 nm starting from an initial fully hybridized duplex. This cutoff was selected as the distance beyond \rood{which} the strands are effectively non-interacting and defines the dissociated state. \rood{We verified that our results were robust to choices of this cutoff over the range 0.8-2.5 nm.} We extracted our computational estimate of $k_d^\mathrm{slow}$ by fitting a decaying exponential to the fraction of hybridized sequences as a function of time $f_\mathrm{hybridized}(t) = \exp(-k_d^\mathrm{slow}t)$. \rood{To correct for finite-size effects associated with the simulation of a single duplex within a periodic simulation box and ensure that we will observe asymptotically correct behavior at long timescales, we apply the fluctuation correction proposed by Ouldridge et al.\citep{Ouldridge2010ExtractingSimulations} to our predictions for $f_\mathrm{hybridized}$ bulk yields. We correct the values observed in our simulations $f_\mathrm{hybridized}^\mathrm{sim}$ to the bulk corrected result $f_\mathrm{hybridized}$ using the following relations:}

@@

\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=\textwidth]{Fig1.pdf}
        \caption{Experimental measurements and computational predictions of slow and fast at T-jump IR responses. Results are reported in terms of the final T-jump temperature. (a) The experimental and simulated slow rate constants $k_d^\mathrm{slow}$ corresponding to duplex dissociation over long time scales. \rood{We apply a fluctuation correction proposed by Ouldridge et al.\citep{Ouldridge2010ExtractingSimulations} to account for finite-size effects and ensure an asymptotically correct dimer fraction} (b) The experimental and simulated fast rate constant $k_d^\mathrm{fast}$ corresponding to terminal base-pair fraying on short time scales. The simulation results are corrected by a sequence-independent scaling factor that corrects for a 10$\times$ acceleration of the slow dissociation dynamics and 120$\times$ acceleration of the fast fraying dynamics. The simulated temperature in all cases is subjected to a (+4) K corrective calibration to account for an observed systematic under-prediction of the melting temperature by the 3SPN.2 model}
        \label{fig:relaxation-comparison}
	\end{center}
\end{figure}

@@

We first compare the thermodynamic predictions for the equilibrium macrostate probabilities \rood{made by the MSM models fitted at the sequence melting temperatures} to those of the nearest neighbor (NN) model as a popular empirical model of DNA hybridization thermodynamics. The NN model predicts the free energy of duplex formation as a sum over helix initiation terms and the hybridization free energies of nearest neighbor pairs of bases that account for both the specific WC pairings and the modulating effects of the local (i.e., nearest neighbor) environment.~\citep{SantaLucia1998AThermodynamics, Santalucia2004TheMotifs} The parameters of the NN model were estimated by regressing over 108 experimental measurements to furnish a predictive model for the free energy of association as a function of DNA sequence and explicitly account for stacking contributions of native pairs, internal mismatches, and dangling ends. We apply the NN model to predict the free energy $F^\mathrm{NN}$ of each of the macrostates occupied by each of the four sequences. A full accounting of our application of the NN model is provided in \blauw{Fig.~S6} of the \blauw{Supporting Information}. The free energy of each macrostate is related to its equilibrium occupancy probability $P$ via the statistical mechanical relationship $F = -k_B T \ln P + C$, where $T$ is temperature, $k_B$ is Boltzmann's constant, and $C$ is an additive constant reflecting our ignorance of the absolute scale of free energies. We use this relationship to convert the equilibrium occupancy probabilities predicted by our MSM and reported in Fig.~\ref{fig:allseq_table}b into free energies $F^\mathrm{MSM}$. The unknown additive constants preclude comparisons of absolute free energies between the MSM and NN model, but it is legitimate to compare relative free energies between pairs of macrostates since the additive constant cancels in taking differences. As such, we arbitrarily set the additive constant $C$ in both the MSM and NN model such that the hybridized state H defines a zero free energy reference state and we report the stability of all macrostates relative to the hybridized state as $\Delta F^\mathrm{NN} = F^\mathrm{NN} - F_H^\mathrm{NN}$ for the NN model and $\Delta F^\mathrm{MSM} = F^\mathrm{MSM} - F_H^\mathrm{MSM}$ for the MSM.

As illustrated in Fig.~\ref{fig:NN_table}, we see that the MSM tends to predict higher free energies for the out-of-register shifted macrostates 5S2, 3S2, 5S4, and 3S4 relative to the H state compared to the NN model. The NN and MSM predictions for the frayed state F4 are in good agreement. Although the shifted state trends are in qualitative agreement, what is the root of the quantitative discrepancy of the MSM and NN models in the predicted relative stabilities? First, the MSM is constructed bottom-up from molecularly detailed 3SPN.2 simulations whereas the NN model is fitted top-down by regression against experimental data. There are approximations inherent in the 3SPN.2 model, not least of which is the coarse-grained representation that integrates over atomic degrees of freedom, and in the NN model that was fitted to limited experimental data assuming a low-order expansion in terms of nearest neighbor additive contributions. Second, although 3SPN.2 is expected to capture some dangling end stabilization effects through base stacking and cross-stacking interactions, the model was not parameterized to fully capture the enthalpic contributions of the interaction of unbound bases with terminal base pairs, whereas this term is explicitly included within the NN model. Third, a well known deficiency of the NN model is the absence of any treatment of inert tails -- free bases that extend beyond the dangling end tend to destabilize the duplex.\citep{DiMichele2014EffectHybridization}

\begin{figure}[ht!]
	\begin{center} 
        \includegraphics[width=160mm, scale=1]{Fig3.pdf}
        \caption{Comparison of the macrostate free energy predictions of the MSMs and nearest neighbor (NN) thermodynamic model \rood{at the sequence melting temperatures}.~\citep{SantaLucia1998AThermodynamics, Santalucia2004TheMotifs} (a) Free energies of each macrostate relative to the hybridized state $\Delta F = F - F_H$. We define the hybridized state H to possess a free energy of zero and take care to only compare relative free energies (i.e., $\Delta F$) between the MSM and NN model. (b) Discrepancy between the macrostate relative free energy predictions $\Delta \Delta F = \Delta F^\mathrm{MSM} - \Delta F^\mathrm{NN}$ of the MSM relative to the NN model. The MSM tends to predict higher relative free energies (i.e., lower occupancy probabilities) for the out-of-register shifted states 5S2, 3S2, 5S4, 3S4 relative to the hybridized state H compared to the NN model.}
        \label{fig:NN_table}
	\end{center}
\end{figure}

@@

In sum, our results demonstrate the profound effect of sequence upon the kinetic landscapes, metastable states, and hybridization/dehybridization mechanisms of short DNA oligomers. Our analysis of this small family of sequences expose preliminary design principles for the (meta)stability of out-of-register and frayed states but we anticipate much greater richness in the landscapes will emerge with studies of longer and more diverse sequences. Going forward, we will extend this work to discern more general trends in sequence-dependent hybridization/dehybridization for a wider range of oligomer sequences and motivate strategies for experimental comparisons. \rood{We also suggest that the MSM approach followed in this work, possibly coupled with biased sampling and reweighting techniques \citep{Wu2014StatisticallyStates, Scherer2015PyEMMAModels}
, may be well-suited to exposing changes in the hybridization/dehybridization mechanism and kinetics as a function of temperature.} We anticipate that these insights may provide foundational design rules by which to improve understanding of \emph{in vivo} hybridization processes and rationally engineer optimized sequences for DNA nanotechnology applications such as DNA-PAINT~\citep{Strauss2020UpDNA-PAINT} and DNA barcoding.~\citep{Shah2019}

>>>>

Over the last couple of decades, DNA has proven to be much more than a vessel for genetic information. From sensing to computing to directed self-assembly, the programmable and predictable nature of DNA has unlocked numerous unforeseen nanotechnology applications.\citep{Seeman2017DNANanotechnology, Adleman1994MolecularProblems, Rothemund2006FoldingPatterns, Gu2010ALine} Recently, single molecule localization techniques have exploited the rapid and transient binding of short DNA oligomers in order to achieve super-resolution microscopy and optical multiplexing. \cite{Schnitzbauer2017Super-resolutionDNA-PAINT, Shah2019, Strauss2020UpDNA-PAINT} Predictive understanding of the sequence-dependent thermodynamics of DNA hybridization/dehybridization -- the assembly/disassembly of a DNA duplex from two single strands -- has underpinned the rational design of DNA oligomer sequences for nanotechnology applications, where sequence-dependent nearest-neighbor models can accurately account for mismatched pairs, dangling ends, and other non-native bonding effects.\citep{SantaLucia1998AThermodynamics, Santalucia2004TheMotifs} Secondary DNA structures such as hairpins and G-quadruplexes have also been studied in depth and leveraged for nanotechnology applications.\citep{Tsukanov2013DetailedOrigami, Mosayebi2014TheFormation, Mergny2019DNANanotechnology} Predictive models of the dynamical, as opposed to purely thermodynamical, behaviors of DNA have become increasingly important in developing technologies such as DNA-PAINT (DNA Points Accumulation For Imaging In Nanoscale Topography), but these technologies have outpaced our fundamental understanding of the dynamics themselves. \citep{Deluca2020DynamicDevices, Cordes2010SensingTransfer, Naimark2020DNADevelopment} Many experimental and computational studies have investigated DNA dynamical phenomena over picosecond to millisecond time scales.\citep{Yin2011KineticsHybridization, Xiao2019, Hinckley2014Coarse-grainedEffects, Sanstead2016, Porschke1971CooperativeTransition} Kinetic models have been developed for particular DNA processes such as toehold exchanges and optical barcoding \citep{Zhang2009ControlExchange, Shah2019ProgrammingFingerprinting} and supervised machine learning techniques have been combined with experimental measurements to predict the on/off rates as a function of sequence. \citep{Schickinger2018TetheredHelices, Zhang2018PredictingSequence, Shah2019}. A comprehensive understanding of the full dynamical landscape of hybridization/dehybridization accounting for the sequence-dependent metastable states and association/dissociation pathways remains lacking and fundamental questions remain unresolved. For example, it remains unclear the extent to which hybridization of short DNA oligomers largely proceeds in a conventionally assumed ``all-or-nothing'' fashion or if long-lived metastable states facilitate the transition.\citep{Araque2016LatticeCooperativity, Sikora2013ModelingIntermediates, Wyer2014KineticsAT-tracts, Sanstead2016, Sanstead2018DirectDehybridization, Xiao2019} Out-of-register ``shifted'' base paired structures \citep{Hinckley2014Coarse-grainedEffects, Maciejczyk2014DNAModel, Araque2016LatticeCooperativity, Xiao2019, Romano2013DNADependence,Flamm2000RNAResolution} and frayed structures \citep{Zgarbova2014BaseRNA, Nonin1995TerminalFraying, Nikolova2012ProbingSimulations, Andreatta2006UltrafastHelix} stand as candidates for metastable states with the potential to mediate substantial deviations from all-or-nothing behavior, but the degree to which these states are kinetically relevant is difficult to determine experimentally and is likely to be highly sequence-dependent. The development of predictive models and design rules with which to engineer DNA strands with tailored hybridization/dehybridization kinetics and pathways is vital to advancing rational design of DNA strands for nanotechnology applications and is also of importance in understanding fundamental biological processes such as transcription and gene regulation.

Our understanding of hybridization dynamics is built upon decades of experiments -- such as temperature-jump, salt-jump, pH-jump, and other perturbative methods -- that rapidly stimulate DNA and monitor relaxation to a new equilibrium.\citep{Morrison1993SensitiveSolution, Wetmur1968KineticsDNA, Craig1971RelaxationOligonucleotides, Porschke1973ThermodynamicsPairs, Williams1989LaserDGCATGC, Narayanan2012ExploringMixing, Chen2007InfluenceHybridization, Sanstead2018DirectDehybridization} More recently, single molecule diffusion and tethered multifluorophore assays have facilitated analyses under equilibrium conditions, but these results can be hampered by slow data collection rates and fluorescent tags effects on strand dynamics, particularly for shorter oligomers.\citep{Liu20173DSolution,  Schickinger2018TetheredHelices, Chen2008Base-by-baseSpectroscopy, Dupuis2013Single-moleculeHelices} A number of computational modeling approaches have also been employed to provide molecular-level resolution of hybridization. Simplified lattice models can recapitulate the essential aspects of the hybridization pathways but lack the realism of continuous space representations \citep{Araque2016LatticeCooperativity, Phys2019}. The long time scales associated with hybridization/dehybridization events place them outside the reach of unbiased all-atom molecular dynamics simulations,\citep{Hinckley2014Coarse-grainedEffects} but they can be observed by employing enhanced sampling techniques\citep{Piana2007AtomisticTransition, Zerze2021ThermodynamicsSimulations, Hinckley2013AnHybridization, Schmitt2013ExploringSurface, Sambriski2009, Sambriski2009SequencePathways, Hoefert2011MolecularOligonucleotides,Romano2013DNADependence, Hinckley2014Coarse-grainedEffects} or by using elevated temperature or denaturing solvent concentrations to induce one-way dissociation events.\citep{Wong2008TheSimulations, Perez2010Real-timeUnfolding} The effect of the applied bias upon the thermodynamics can be rigorously corrected for using standard reweighting techniques \citep{Gallicchio2005TemperaturePaths, Souaille2001ExtensionCalculations, Shirts2008StatisticallyStates, Kumar1992THEMethod}. \rood{Rigorous elimination of the bias in the kinetics is critical for the construction of robust and reliable Markov state models that reflect the true system dynamics and are uncontaminated by any residual effects of the biasing potentials used to induce good sampling. A number of approaches to correct the kinetics are also available, including Girsanov reweighting, transition-based reweighting analysis (TRAM), dynamic histogram analysis method (DHAM), and their derivatives~\citep{Prinz2011OptimalDynamics, Chodera2011DynamicalTemperatures, Stelzl2017DynamicSimulations, Donati2017GirsanovModels, Donati2018GirsanovSimulations, Quer2018ANCOORDINATES} \citep{Mey2014XTRAM:States}. The application of these methods under the conditions of high bias necessary for good sampling can, however, present challenges for numerical convergence.} A number of coarse-grained DNA force fields have been developed that enable direct observation of these events over microsecond time scales via unbiased coarse-grained molecular dynamics simulations, \citep{Romano2013DNADependence, Hinckley2013AnHybridization, Maciejczyk2014DNAModel, Markegard2015, Dans2016MultiscaleDNA} which, up to an acceleration factor associated with the smoothing of the underlying free energy landscape inherent to the coarse-graining procedure, can preserve a faithful model of the unbiased dynamics and associated pathways. These models have previously been used to study biological phenomena such as nucleosome dynamics\citep{Lequieu2016Tension-dependentUnwrapping, Lequieu2017InSliding} and transcription factor binding \cite{Terakawa2015P53Simulations, Tan2018DynamicData} as well as nanotechnology applications such as strand displacement \citep{Srinivas2013OnDisplacement, Haley2020DesignDisplacement} and DNA origami \citep{Snodin2019Coarse-grainedOrigami, Doye2020TheOrigami}. \rood{In this study, we choose to employ a coarse-grained model for DNA \citep{Hinckley2013AnHybridization} that is sufficiently inexpensive to enable the collection of sufficient volumes of unbiased simulation trajectories and adequately sample configurational space that we do not need to appeal to biasing strategies to enhance convergence nor apply any \textit{post hoc} corrections to the thermodynamics or kinetics.}

@@

We first sought to demonstrate that the coarse-grained 3SPN.2 model accurately recapitulates the experimentally observed kinetics of DNA oligomer hybridization/dissociation by validating our computational predictions against temperature-jump infrared (T-jump IR) experiments. We conducted T-jump IR experiments for each DNA sequence as a function of temperature and extracted \rood{the ``slow'' $k_d^\mathrm{slow}$  and ``fast'' $k_d^\mathrm{fast}$ rates, corresponding to processes proceeding on 10-30 $\mu$s and 70-100 ns timescales, respectively}. The slow response has previously been attributed to duplex dissociation on microsecond time scales induced by the rapid heating of the initially hybridized duplex.~\citep{Sanstead2016, Sanstead2018DirectDehybridization} The fast response has been assigned to terminal base pair fraying,~\citep{Sanstead2016, Sanstead2018DirectDehybridization} with this process corresponding to a relatively complex dynamical process that can span time scales from picoseconds to microseconds.~\cite{Nonin1995TerminalFraying, Nikolova2012ProbingSimulations, Andreatta2006UltrafastHelix, Galindo-Murillo2015ConvergenceDGCACGAACGAACGAACGC} All-atom simulations suggest that frayed ends can assume misaligned WC bonds, base-sugar hydrogen bonds, and terminal stacked conformations.~\citep{Pinamonti2019TheModels, Zgarbova2014BaseRNA}

To computationally mimic the T-jump process in our 3SPN.2 simulations, we conducted 1 $\mu$s simulations of an initially hybridized DNA duplex over a range of temperatures and monitored its structural relaxation. We performed 120 independent simulations for each DNA sequence at each temperature, and from these extracted computational estimates of $k_d^\mathrm{slow}$ and $k_d^\mathrm{fast}$ (\blauw{Fig.~S4}). First, we tracked the slow response corresponding to duplex dissociation in our simulations by compiling the distribution of times at which both of the central base pairs first separate to a distance of 1.3 nm starting from an initial fully hybridized duplex. This cutoff was selected as the distance beyond \rood{which} the strands are effectively non-interacting and defines the dissociated state. \rood{We verified that our results were robust to choices of this cutoff over the range 0.8-2.5 nm.} We extracted our computational estimate of $k_d^\mathrm{slow}$ by fitting a decaying exponential to the fraction of hybridized sequences as a function of time $f_\mathrm{hybridized}(t) = \exp(-k_d^\mathrm{slow}t)$. \rood{To correct for finite-size effects associated with the simulation of a single duplex within a periodic simulation box and ensure that we will observe asymptotically correct behavior at long times, we apply the fluctuation correction proposed by Ouldridge et al.\citep{Ouldridge2010ExtractingSimulations} to our predictions for $f_\mathrm{hybridized}$ bulk yields. We correct the values observed in our simulations $f_\mathrm{hybridized}^\mathrm{sim}$ to the bulk corrected result $f_\mathrm{hybridized}$ as,}

@@

\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=\textwidth]{Fig1.pdf}
        \caption{Experimental measurements and computational predictions of slow and fast at T-jump IR responses. Results are reported in terms of the final T-jump temperature. (a) The experimental and simulated slow rate constants $k_d^\mathrm{slow}$ corresponding to duplex dissociation over long time scales. \rood{We apply a fluctuation correction proposed by Ouldridge et al.\citep{Ouldridge2010ExtractingSimulations} to account for finite-size effects and ensure an asymptotically correct dimer fraction at long times.} (b) The experimental and simulated fast rate constant $k_d^\mathrm{fast}$ corresponding to terminal base-pair fraying on short time scales. The simulation results are corrected by a sequence-independent scaling factor that corrects for a 10$\times$ acceleration of the slow dissociation dynamics and 120$\times$ acceleration of the fast fraying dynamics. The simulated temperature in all cases is subjected to a (+4) K corrective calibration to account for an observed systematic under-prediction of the melting temperature by the 3SPN.2 model}
        \label{fig:relaxation-comparison}
	\end{center}
\end{figure}

@@

We first compare the thermodynamic predictions for the equilibrium macrostate probabilities \rood{made by the MSM models fitted at the sequence melting temperatures} to those of the nearest neighbor (NN) model as a popular empirical model of DNA hybridization thermodynamics. The NN model predicts the free energy of duplex formation as a sum over helix initiation terms and the hybridization free energies of nearest neighbor pairs of bases that account for both the specific WC pairings and the modulating effects of the local (i.e., nearest neighbor) environment.~\citep{SantaLucia1998AThermodynamics, Santalucia2004TheMotifs} The parameters of the NN model were estimated by regressing over 108 experimental measurements to furnish a predictive model for the free energy of association as a function of DNA sequence and explicitly account for stacking contributions of native pairs, internal mismatches, and dangling ends. We apply the NN model to predict the free energy $F^\mathrm{NN}$ of each of the macrostates occupied by each of the four sequences. A full accounting of our application of the NN model is provided in \blauw{Fig.~S6} \rood{and supporting text within} the \blauw{Supporting Information}. The free energy of each macrostate is related to its equilibrium occupancy probability $P$ via the statistical mechanical relationship $F = -k_B T \ln P + C$, where $T$ is temperature, $k_B$ is Boltzmann's constant, and $C$ is an additive constant reflecting our ignorance of the absolute scale of free energies. We use this relationship to convert the equilibrium occupancy probabilities predicted by our MSM and reported in Fig.~\ref{fig:allseq_table}b into free energies $F^\mathrm{MSM}$. The unknown additive constants preclude comparisons of absolute free energies between the MSM and NN model, but it is legitimate to compare relative free energies between pairs of macrostates since the additive constant cancels in taking differences. As such, we arbitrarily set the additive constant $C$ in both the MSM and NN model such that the hybridized state H defines a zero free energy reference state and we report the stability of all macrostates relative to the hybridized state as $\Delta F^\mathrm{NN} = F^\mathrm{NN} - F_H^\mathrm{NN}$ for the NN model and $\Delta F^\mathrm{MSM} = F^\mathrm{MSM} - F_H^\mathrm{MSM}$ for the MSM.

As illustrated in Fig.~\ref{fig:NN_table}, we see that the MSM tends to predict higher free energies for \rood{all macrostates relative to the H state compared to the NN model}. Although the trends are in qualitative agreement, what is the root of the quantitative discrepancy of the MSM and NN models in the predicted relative stabilities? First, the MSM is constructed bottom-up from molecularly detailed 3SPN.2 simulations whereas the NN model is fitted top-down by regression against experimental data. There are approximations inherent in the 3SPN.2 model, not least of which is the coarse-grained representation that integrates over atomic degrees of freedom, and in the NN model that was fitted to limited experimental data assuming a low-order expansion in terms of nearest neighbor additive contributions. Second, although 3SPN.2 is expected to capture some dangling end stabilization effects through base stacking and cross-stacking interactions, the model was not parameterized to fully capture the enthalpic contributions of the interaction of unbound bases with terminal base pairs, whereas this term is explicitly included within the NN model. Third, a well known deficiency of the NN model is the absence of any treatment of inert tails -- free bases that extend beyond the dangling end tend to destabilize the duplex.\citep{DiMichele2014EffectHybridization}

\begin{figure}[ht!]
	\begin{center} 
        \includegraphics[width=160mm, scale=1]{Fig3.pdf}
        \caption{Comparison of the macrostate free energy predictions of the MSMs and nearest neighbor (NN) thermodynamic model \rood{at the sequence melting temperatures}.~\citep{SantaLucia1998AThermodynamics, Santalucia2004TheMotifs} (a) Free energies of each macrostate relative to the hybridized state $\Delta F = F - F_H$. We define the hybridized state H to possess a free energy of zero and take care to only compare relative free energies (i.e., $\Delta F$) between the MSM and NN model. (b) Discrepancy between the macrostate relative free energy predictions $\Delta \Delta F = \Delta F^\mathrm{MSM} - \Delta F^\mathrm{NN}$ of the MSM relative to the NN model. The MSM tends to predict higher relative free energies (i.e., lower occupancy probabilities) relative to the hybridized state H compared to the NN model.}
        \label{fig:NN_table}
	\end{center}
\end{figure}

@@

In sum, our results demonstrate the profound effect of sequence upon the kinetic landscapes, metastable states, and hybridization/dehybridization mechanisms of short DNA oligomers. Our analysis of this small family of sequences expose preliminary design principles for the (meta)stability of out-of-register and frayed states but we anticipate much greater richness in the landscapes will emerge with studies of longer and more diverse sequences. Going forward, we will extend this work to discern more general trends in sequence-dependent hybridization/dehybridization for a wider range of oligomer sequences and motivate strategies for experimental comparisons. \rood{We also suggest that the MSM approach followed in this work, possibly coupled with biased sampling and reweighting techniques \citep{Wu2014StatisticallyStates, Scherer2015PyEMMAModels}, may be well-suited to expose changes in the hybridization/dehybridization mechanism and kinetics as a function of temperature.} We anticipate that these insights may provide foundational design rules by which to improve understanding of \emph{in vivo} hybridization processes and rationally engineer optimized sequences for DNA nanotechnology applications such as DNA-PAINT~\citep{Strauss2020UpDNA-PAINT} and DNA barcoding.~\citep{Shah2019}
